---
title: "p8105_hw1_mk5149"
author: "Mungyu Kwok"
date: "2025-09-16"
output: github_document
---

This is the R markdown file for Homework 1, P8105 Data Science, created by Mungyu Kwok on September 16th 2025.

Firstly, we load any necessary library that we will use in this file.
```{r}
library(tidyverse)
library(moderndive)
```

# Problem 1
We start problem 1 by loading the data from the `early_january_weather`.
```{r}
# Load the data
data("early_january_weather")
```

Next, we extract the essential information about the dataset: variable names, number of wors & columns, and compute the mean temperature.
```{r}
# Get the information about the dataset
names(early_january_weather)   # variable names
nrow(early_january_weather)    # number of rows
ncol(early_january_weather)    # number of columns
mean(early_january_weather$temp, na.rm = TRUE)   # mean temperature
```
The `early_january_weather` dataset contains `r nrow(early_january_weather)` rows and 
`r ncol(early_january_weather)` columns.  
Important variables include `origin`, `year`, `month`, `day`, `hour`, `temp`, `dewp`, `humid`, `wind_dir`, `wind_dir`, `wind_speed`, `wind_gust`, `precip`, `pressure`, `visib` and `time_hour`.  
The mean temperature during this period was `r mean(early_january_weather$temp, na.rm = TRUE)` degrees Fahrenheit.

Based on the information obtained, we compute a scatterplot (time versus temperature).
```{r}
# Generate the scatter plot
scatter_temp_time <- ggplot(early_january_weather, 
                           aes(x = time_hour, y = temp, color = humid)) + geom_point(size = 2) + 
  labs(x = "Time (hour)", y = "Temperature", color = "Humidity") + theme_minimal()
scatter_temp_time

ggsave("scatter_temp_vs_time.png", scatter_temp_time, width = 7, height = 4)
```
The scatterplot shows the relationship between time (x-axis) and temperature (y-axis)
in early January, with point color indicating humidity. Temperatures generally
increased over the course of the two weeks, although fluctuations occurred within
each day. Higher humidity values (lighter blue points) appear more often during
periods of higher temperature, especially in the second week of January.
Overall, the plot suggests a warming trend accompanied by varying levels of humidity.


# Problem 2
We first create 4 vectors:
a numeric vector (10 random samples from the standard normal distribution),
a logicala vector indicating whether elements of the sample are greater than 0,
a character vector of length 10,
a factor vector of length 10, with 3 different factor “levels”.
```{r}
# Create the numeric vector
numeric_vec = rnorm(10)

# Create the logical vector, TRUE if numeric_vec > 0
logical_vec = numeric_vec > 0

# Create the character vector
character_vec = c("a","b","c","d","e","f","g","h","i","j")

# Create the factor vector
factor_vec = factor(c("low","medium","high","low","medium","high","low","medium","high","low"))
```

Then, we combine them into a dataframe.
```{r}
df = tibble(
  numeric_vec,
  logical_vec,
  character_vec,
  factor_vec
)
df
```

Next, we attempt to calculate the mean of each variable.
```{r}
mean(pull(df, numeric_vec))     # works, numeric mean
mean(pull(df, logical_vec))     # works, TRUE = 1 and FALSE = 0
mean(pull(df, character_vec))   # error, characters cannot be averaged
mean(pull(df, factor_vec))      # error, factors are not directly numeric
```
We can see that, the mean of the numeric vector works as expected. The mean of the logical vector also works, because R treats TRUE as 1 and FALSE as 0. However, the mean of the character and factor variables does not work, because they are numeric. 








