---
title: "p8105_hw1_mk5149"
author: "Mungyu Kwok"
date: "2025-09-16"
output: github_document
---

This is the R markdown file for Homework 1, P8105 Data Science, created by Mungyu Kwok on September 16th 2025.

```{r}
library(tidyverse)
library(moderndive)
```

# Problem 1
```{r}
# Load the data
data("early_january_weather")
```

```{r}
# Get the information about the dataset
names(early_january_weather)   # variable names
nrow(early_january_weather)    # number of rows
ncol(early_january_weather)    # number of columns
mean(early_january_weather$temp, na.rm = TRUE)   # mean temperature
```
The `early_january_weather` dataset contains `r nrow(early_january_weather)` rows and 
`r ncol(early_january_weather)` columns.  
Important variables include `origin`, `year`, `month`, `day`, `hour`, `temp`, `dewp`, `humid`, `wind_dir`, `wind_dir`, `wind_speed`, `wind_gust`, `precip`, `pressure`, `visib` and `time_hour`.  
The mean temperature during this period was `r mean(early_january_weather$temp, na.rm = TRUE)` degrees Fahrenheit.

```{r}
# Generate the scatter plot
scatter_temp_time <- ggplot(early_january_weather, 
                           aes(x = time_hour, y = temp, color = humid)) + geom_point(size = 2) + 
  labs(x = "Time (hour)", y = "Temperature", color = "Humidity") + theme_minimal()
scatter_temp_time

ggsave("scatter_temp_vs_time.png", scatter_temp_time, width = 7, height = 4)
```
The scatterplot shows the relationship between time (x-axis) and temperature (y-axis)
in early January, with point color indicating humidity. Temperatures generally
increased over the course of the two weeks, although fluctuations occurred within
each day. Higher humidity values (lighter blue points) appear more often during
periods of higher temperature, especially in the second week of January.
Overall, the plot suggests a warming trend accompanied by varying levels of humidity.



